---
title: "R Notebook"
output: html_notebook

purpose: replace USA with states and add additional rows for new simUIDs in SimUIDLUIDHRU_map in SimUIDlu_map.gms
---

```{r}
library(sf)
library(GISTools)
library(rgdal)
require(ggplot2)
library(dplyr)
```
=============================================================
=============================================================
# A. Load non GAMS inputs
=============================================================
## Read simU shapefile
```{r}
simU_dir <- "../../SimU/gis/"
simU_fn_new <- "SimU_all_select840_USstateSplit_proj_newSUcol_v2.shp"
simU_new <- st_read(paste(simU_dir, simU_fn_new, sep=""))
colnames(simU_new)
## NAME_1 = state names
## SimuIDn = new simulation unit ID

## subset to select columns
simU_new_sub <- simU_new %>%
  select(c(NAME_1, SimUID, SimUIDn), STATE = NAME_1)

unique(simU_new_sub$STATE)
```
=============================================================
## read state abbreviations
```{r}
abbv <- read.csv("inputs/stateAbbrev.csv", header=TRUE)
colnames(abbv) <- c("STATE", "shorthand", "STATE_Abbrev")

## join to SimU data frame and convert to nonspatial dataframe
simU_new_sub_abbrev <- simU_new_sub %>% 
  as.data.frame() %>%
  select(-geometry) %>%
  right_join(abbv, by = c("STATE" = "STATE"))

## check join
filter(simU_new_sub_abbrev, is.na(STATE_Abbrev))
```
=============================================================
=============================================================
# B. UPDATE GAMS map: SimUIDLUIDHRU_map in Model/SimUIDl_map.gms
Requires you to run part A

## read gams map as txt file
```{r}
mapping <- read.csv("inputs/SimUIDLUIDHRU_map.txt", sep = ".", header=FALSE)
colnames(mapping) <- c("SimUID", "Country", "LU", "alt", "slope", "soil")
```

## join with simUID vector attribute table and write to csv
```{r}
## join GAMS map to SimUID attribute table using outer join
mapping.join <- mapping %>%
  full_join(simU_new_sub_abbrev, by = c("SimUID" = "SimUID")) %>%
  ## Country names have trailing white spaces, use trimw function to remove trailing 
  mutate(Country = trimws(Country, "right")) %>%
  mutate(alt = trimws(alt, "right")) #%>%
  #filter(Country == "USA")

## replace SimUID column with SimUIDn column if it's USA
## replace Country column with STATE column if it's USA
mapping.join.USA <- mapping.join %>%
  mutate(SimUID = if_else(Country == "USA", as.character(SimUIDn), as.character(SimUID))) %>%
  mutate(Country = if_else(Country == "USA", as.character(STATE_Abbrev), as.character(Country)))

## check to see if there are any unmatched rows 
filter(mapping.join.USA, is.na(alt))

################
## write to csv, use . as separator for SimUIDLUIDHRU_map.txt
## drop the STATE and SimUIDn variables/columns
mapping.join.USA.final <- mapping.join.USA %>% select(-c(STATE, SimUIDn, shorthand, STATE_Abbrev))

## write to CSV
write.table(mapping.join.USA.final, "outputs/SimUIDLUIDHRU_map.txt", col.names = FALSE, row.names = FALSE, sep = ". \t", quote=FALSE)
```

=============================================================
=============================================================
# C. Update SimUIDLUID_map in sets_colrow.gms
Requires you to run part A and B
```{r}
## drop the STATE and SimUIDn variables/columns
mapping.join.USA.final.SimUIDLUID_map <- mapping.join.USA %>% 
  select(c(SimUID, LU)) %>%
  distinct()

## write to csv, use . as separator for SimUIDLUID_map
write.table(mapping.join.USA.final.SimUIDLUID_map, "outputs/SimUIDLUID_map.txt", col.names = FALSE, row.names = FALSE, sep = ". \t", quote=FALSE)
```

=============================================================
=============================================================
# D. CREATE NEW SimUID to state map: map_US_SIMUID_state in sets_US.gms
Requires you to run part A and B
```{r}
mapping.join.USA.final.SimUIDstate_map <- mapping.join.USA %>% 
  select(c(SimUID, Country)) %>%
  distinct() %>%
  filter(Country %in% abbv$STATE_Abbrev)

write.table(mapping.join.USA.final.SimUIDstate_map, "outputs/map_US_SIMUID_state.txt", col.names = FALSE, row.names = FALSE, sep = ".\t", quote=FALSE)
```

=============================================================
=============================================================
# E. Update GAMS map: SimUID_MAP(SimUID,ALLCOUNTRY,ColRow30,AltiClass,SlpClass,SoilClass) in Data/SimUID_map.gms
Requires you to run part A
```{r}
mapping <- read.csv("inputs/SimUID_MAP.txt", sep = ".", header=FALSE)
colnames(mapping) <- c("SimUID", "ALLCOUNTRY", "ColRow30", "AltiClass", "SlpClass", "SoilClass")
```

## join with simUID vector attribute table and write to csv
```{r}
## join GAMS map to SimUID attribute table using outer join and trim trailing white space
mapping.join <- mapping %>%
  full_join(simU_new_sub_abbrev, by = c("SimUID" = "SimUID")) %>%
  ## Country names have trailing white spaces, use trimw function to remove trailing 
  mutate(ALLCOUNTRY = trimws(ALLCOUNTRY, "right")) %>%
  mutate(AltiClass = trimws(AltiClass, "right")) #%>%
  #filter(Country == "USA")

## replace SimUID column with SimUIDn column if it's USA
## replace Country column with STATE column if it's USA
mapping.join.USA <- mapping.join %>%
  mutate(SimUID = if_else(ALLCOUNTRY == "USA", as.character(SimUIDn), as.character(SimUID))) %>%
  mutate(ALLCOUNTRY = if_else(ALLCOUNTRY == "USA", as.character(STATE_Abbrev), as.character(ALLCOUNTRY)))

## check to see if there are any unmatched rows 
filter(mapping.join.USA, is.na(AltiClass))

################
## write to csv, use . as separator for SimUID_MAP.txt

## drop the STATE and SimUIDn variables/columns
mapping.join.USA.final <- mapping.join.USA %>% select(-c(STATE, SimUIDn, shorthand, STATE_Abbrev))

## write to CSV
write.table(mapping.join.USA.final, "outputs/SimUID_MAP.txt", col.names = FALSE, row.names = FALSE, sep = ". \t", quote=FALSE)
```

=============================================================
=============================================================
## F. CREATE NEW state abbrev: US_states(ANYREGION) in sets_US.gms
Requires you to run part A
```{r}
abbv
write.table(abbv$STATE_Abbrev, "outputs/stateAbbrevList.txt", col.names = FALSE, row.names = FALSE, sep = ",", quote=FALSE, eol = ",")
```

